Aspects of Dealing with Imperfect Data in Temporal Databases

### 2012-03-13 book/author.tex
Almost done! In this one, the references are a little improved and the piece about fuzzy temporal reasoning (without databases) is written out.

\title*{An Overview on Temporal Databases.\\
 New Approaches and Proposals.}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Jos\'e Pons \and Christophe Billiet \and Olga Pons \and Guy De Tr\'e}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{
Jos\'e Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
\email{jpons@decsai.ugr.es}
\and Christophe Billiet \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Christophe.Billiet@telin.ugent.be}
\and Olga Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
\email{opc@decsai.ugr.es}
\and Guy De Tr\'e \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Guy.DeTre@telin.ugent.be}
}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

\vspace{-35pt}
%\abstract*{There exist some objects in the nature that have a time-variant (e.g. an employee's working life) or time-related (e.g. the transactions in a bank account) nature. Modelling these objects in a (relational) database is possible, but the time-related attributes have an impact on the consistency of the entire database. Therefore, some temporal database models have been proposed to deal with this. These models deal both with the problem of representation and querying. Time itself is a source of imprecision / vagueness / uncertainty since the measures of time are inherently imprecise. Humans beings manage the time with imprecision and vagueness. Nevertheless, the amount of uncertainty in the temporal indications are defined by socially well-known boundaries. Several proposals for dealing with imprecision in time have been done: some proposals consider the basis of the system the imprecision related to the conversion of diferent time points and intervals. While some other proposals consider the specification of the time as the source of imprecision: the time points are ill-known.}

\abstract*{
In reality, some objects or concepts have properties with a time-variant (e.g. an employee's job function) or time-related (e.g. financial transactions of a bank account) nature. Modelling these kinds of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database. Therefore, some temporal database models have been proposed to deal with this. Time itself is a source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect. Accordingly, human beings manage time using temporal indications and temporal notions, which may contain imprecision, vagueness and uncertainty. However, the imperfection in human-used temporal indications is to some degree constrained by socially well-known boundaries. Several proposals for dealing with such imperfections exist. Some of these proposals consider the basis of the system to be the conversion of the specificity of temporal notions between used temporal expressions. Other proposals consider the temporal indications in the used temporal expressions to be the source of imperfection. In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that arise during or as a result of this modelling. Next to this, a novel and currently researched technique for handling some of these imperfections is presented.
}

\abstract{
In reality, some objects or concepts have properties with a time-variant (e.g. an employee's job function) or time-related (e.g. financial transactions of a bank account) nature. Modelling these kinds of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database. Therefore, some temporal database models have been proposed to deal with this. Time itself is a source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect. Accordingly, human beings manage time using temporal indications and temporal notions, which may contain imprecision, vagueness and uncertainty. However, the imperfection in human-used temporal indications is to some degree constrained by socially well-known boundaries. Several proposals for dealing with such imperfections exist. Some of these proposals consider the basis of the system to be the conversion of the specificity of temporal notions between used temporal expressions. Other proposals consider the temporal indications in the used temporal expressions to be the source of imperfection. In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that arise during or as a result of this modelling. Next to this, a novel and currently researched technique for handling some of these imperfections is presented.
}

@@

\section{Time Modelling}
\label{sec:time-domain}
\input{sources/timeDomain.tex}
\def\JPicScale{1}
\section{Temporal Databases}
\label{sec:temporal-databases}
\input{sources/temporalDatabases.tex}
\def\JPicScale{1}
\section{Imperfection in Temporal Databases}
\label{sec:imprecision-uncertainty}
\input{sources/fuzzyTemporalDatabases.tex}

%\section{Bipolarity on temporal databases}
%\label{sec:bipolarity}
%\input{sources/bipolarity.tex}
\def\JPicScale{1}
\section{Conclusions and Further Research Work}
\label{sec:further-research-work}
\input{sources/frw.tex}
\def\JPicScale{1}
\section*{Acknowledgements}
Part of this research is supported by the grant BES-2009-013805 within the research project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in relational DBMS}.

>>>>

\title*{Aspects of Dealing with Imperfect Data in Temporal Databases}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Jos\'e Pons \and Christophe Billiet \and Olga Pons \and Guy De Tr\'e}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
%\institute{
%Jos\'e Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
%\email{jpons@decsai.ugr.es}
%\and Christophe Billiet \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Christophe.Billiet@telin.ugent.be}
%\and Olga Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
%\email{opc@decsai.ugr.es}
%\and Guy De Tr\'e \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Guy.DeTre@telin.ugent.be}
%}

\institute{
Jos\'e Pons, Olga Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
\email{jpons, opc@decsai.ugr.es}
\and
Christophe Billiet, Guy De Tr\'e \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Christophe.Billiet, Guy.DeTre@telin.ugent.be}
}

%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

\vspace{-70pt}
%\abstract*{There exist some objects in the nature that have a time-variant (e.g. an employee's working life) or time-related (e.g. the transactions in a bank account) nature. Modelling these objects in a (relational) database is possible, but the time-related attributes have an impact on the consistency of the entire database. Therefore, some temporal database models have been proposed to deal with this. These models deal both with the problem of representation and querying. Time itself is a source of imprecision / vagueness / uncertainty since the measures of time are inherently imprecise. Humans beings manage the time with imprecision and vagueness. Nevertheless, the amount of uncertainty in the temporal indications are defined by socially well-known boundaries. Several proposals for dealing with imprecision in time have been done: some proposals consider the basis of the system the imprecision related to the conversion of diferent time points and intervals. While some other proposals consider the specification of the time as the source of imprecision: the time points are ill-known.}

\abstract*{
In reality, some objects or concepts have properties with a time-variant or time-related nature. Modelling these kinds of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database. Therefore, temporal database models have been proposed to deal with this. Time itself can be at the source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect. Accordingly, human beings manage time using temporal indications and temporal notions, which may contain imprecision, vagueness and uncertainty. However, the imperfection in human-used temporal indications is supported by human interpretation, whereas information systems need extraordinary support for this. Several proposals for dealing with such imperfections when modelling temporal aspects exist. Some of these proposals consider the basis of the system to be the conversion of the specificity of temporal notions between used temporal expressions. Other proposals consider the temporal indications in the used temporal expressions to be the source of imperfection. In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that may arise during or as a result of this modelling. Next to this, a novel and currently researched technique for handling some of these imperfections is presented.
}

\abstract{
In reality, some objects or concepts have properties with a time-variant or time-related nature. Modelling these kinds of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database. Therefore, temporal database models have been proposed to deal with this. Time itself can be at the source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect. Accordingly, human beings manage time using temporal indications and temporal notions, which may contain imprecision, vagueness and uncertainty. However, the imperfection in human-used temporal indications is supported by human interpretation, whereas information systems need extraordinary support for this. Several proposals for dealing with such imperfections when modelling temporal aspects exist. Some of these proposals consider the basis of the system to be the conversion of the specificity of temporal notions between used temporal expressions. Other proposals consider the temporal indications in the used temporal expressions to be the source of imperfection. In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that may arise during or as a result of this modelling. Next to this, a novel and currently researched technique for handling some of these imperfections is presented.
}

@@

\section{Basic Concepts and Issues in Time Modelling}
\label{sec:time-domain}
\input{sources/timeDomain.tex}
\def\JPicScale{1}

\section{Data Imperfections in Time Modelling}
\label{sec:fuzzy-time-mod}
\input{sources/fuzzyTimeDomain.tex}

\section{Basic Concepts and Issues in Temporal Databases}
\label{sec:temporal-databases}
\input{sources/temporalDatabases.tex}
\def\JPicScale{1}

\section{Data Imperfections in Temporal Databases}
\label{sec:imprecision-uncertainty}
\input{sources/fuzzyTemporalDatabases.tex}

%\section{Bipolarity on temporal databases}
%\label{sec:bipolarity}
%\input{sources/bipolarity.tex}
\def\JPicScale{1}
\section{Conclusions and Further Research}
\label{sec:further-research-work}
\input{sources/frw.tex}
\def\JPicScale{1}
\section*{Acknowledgements}
Part of this research is supported by the grant BES-2009-013805 within the research project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in relational DBMS}.

### 2012-03-13 book/sources/intro.tex
Almost done! In this one, the references are a little improved and the piece about fuzzy temporal reasoning (without databases) is written out.

A very specific type of information systems are database systems, which are computer systems designed to manage databases. A database contains data representing real objects or concepts. Every one of these data is a result value of a measurement of a property or a description of a property of a real object or concept. In reality, some aspects or properties of objects or concepts are time-variant or time-related. E.g. the moment of a bank transaction is traditionally a moment in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant. A temporal database schema is a database schema that models real objects or concepts with time-related or time-variant properties. However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints. An example. Consider a relation in a relational library database, modelling the presence of books in the library. Every physical book is represented by a unique identifier. Every record in the relation contains such an identifier, a date on which the corresponding book was loaned and a date on which it was subsequently returned (if it was returned). Without further precautions, a library employee could add several records with the same book identifier, different `loaned'-dates and no `returned'-dates. This group of records would represent the same physical book being loaned several times on different dates and never returned, which is of course impossible. A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.

@@

An interesting issue in temporal modelling concerns relationships between temporal notions. Notably, Allen~\cite{Allen83} studied temporal relationships between time intervals (and as a special case time points). Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed for richer and complexer user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query and the temporal indications in the database.

Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they are often not hindered by imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in the descriptions of these temporal notions. Among many others, these possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling. E.g. the description of the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 1/01/1201 A.D. and 31/12/1300 A.D.' contains imperfection because of the uncertainty in the used time-related expression. It is known that the building was finished on a single day, but it is not known precisely which day this was.

To allow information systems to cope with these and similar imperfections, many approaches adopt fuzzy sets~\cite{Zadeh65} for the representation of temporal information~\cite{343607},~\cite{nagypal2003},~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy},~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in}. The temporal relationships studied by Allen were fuzzified by several authors~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08}. Garrido et al. ~\cite{garrido2009} present different temporal operators, defined by a combination of regular fuzzy comparisons. Both~\cite{garrido2009} and~\cite{Pon11} deal with uncertainty in temporal expressions concerning time intervals. Other approaches, like~\cite{Qia09}, use rough sets~\cite{Pawlak1995} to represent time intervals.

@@

The rest of this chapter is structured as follows. Section \ref{sec:time-domain} presents some basic concepts and terminology about temporal modelling and discusses some of its important aspects and issues. In section \ref{sec:temporal-databases}, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems. Finally, in section \ref{sec:imprecision-uncertainty}, an approach to querying temporal databases containing imperfect temporal information is presented, followed by some conclusions and some suggestions for future work in section \ref{sec:further-research-work}.

>>>>

A very specific type of information systems are database systems, which are computer systems designed to manage databases. A database contains data representing real objects or concepts. Each (atomic) part of these data is a result value of a measurement of a property or a description of a property of a real object or concept. In reality, some aspects or properties of objects or concepts are time-variant or time-related. E.g., the moment of a bank transaction is traditionally a moment in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant. A temporal database schema is a database schema that models real objects or concepts with time-related or time-variant properties. However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints. An example. Consider a relation in a relational library database, modelling the presence of books in the library. Every physical book is represented by a unique identifier. Every record in the relation contains such an identifier, a date on which the corresponding book was loaned and a date on which it was subsequently returned (if it was returned). Without further precautions, a library employee could add several records with the same book identifier, different `loaned'-dates and no `returned'-dates. This group of records would represent the same physical book being loaned several times on different dates and never returned, which is of course impossible. A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.

@@

An interesting issue in temporal modelling concerns relationships between temporal notions. Notably, Allen~\cite{Allen83} studied temporal relationships between time intervals (and as a special case time points). Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed for richer and more complex user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query and the temporal indications in the database. E.g., in a relation modelling who was department head of an institution during which periods of time, a query like `who were the department heads when Thomas worked for the institution' can be evaluated using similar relationships.

Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they often have to deal with imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in the descriptions of these temporal notions. Among many others, these possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling. E.g., the description of the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D. and 31/12/1300 A.D.' contains imperfection because of the uncertainty in the used time-related expression. It is known that the building was finished on a single day, but it is not known precisely which day this was.

To allow information systems to cope with these and similar data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh65} for the representation of temporal information~\cite{343607},~\cite{nagypal2003},~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy},~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in}. The temporal relationships studied by Allen were fuzzified by several authors~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08}. Garrido et al. ~\cite{garrido2009} present different temporal operators, defined by a combination of regular fuzzy comparisons. Both~\cite{garrido2009} and~\cite{Pon11} deal with uncertainty in temporal expressions concerning time intervals. Other approaches, like~\cite{Qia09}, use rough sets~\cite{Pawlak1995} to represent time intervals.

@@

The aim of this chapter is to present and explain some main concepts regarding time in information systems and to present and discuss some issues and techniques concerned with handling data imperfections related to time. The rest of this chapter is structured as follows. Section \ref{sec:time-domain} presents some basic concepts and terminology about temporal modelling and discusses some of its important aspects and issues, while section~\ref{sec:fuzzy-time-mod} presents some important issues concerning the combination of data imperfections and temporal modelling. In section \ref{sec:temporal-databases}, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems. Finally, in section \ref{sec:imprecision-uncertainty}, an approach to querying temporal databases containing imperfect temporal information is presented, followed by some conclusions and some suggestions for future work in section \ref{sec:further-research-work}.

### 2012-03-13 book/sources/timeDomain.tex
Almost done! In this one, the references are a little improved and the piece about fuzzy temporal reasoning (without databases) is written out.

Data models used by information systems (and in specific, temporal database systems) may represent these underlying time domains using \emph{chronons}~\cite{Dyreson1994}. However, to explain what chronons are, an explanation of some other temporal concepts is necessary.

@@

\begin{definition}\textbf{Temporal Element}\\
A \emph{\textbf{temporal element}} is a finite union of time intervals.
\end{definition}

@@

\begin{definition}\textbf{Duration}\\
A \emph{\textbf{duration}} is an amount of time with known length, but no specific starting or ending instants.
\end{definition}

@@

Data models might now represent an underlying time domain using chronons:

@@

In a system, the granularity with the shortest granules is the \emph{chronons granularity}, which is denoted as `$\bot$'. It is the granularity of which the granules are chronons. 
%It is now possible to define a function $f$, called a \emph{mapping function}, that maps a given granule $i_G$, $i \in \Z$, from a given granularity $G$ to a set of corresponding chronons. The function definition should look like this:

@@

\begin{align}
f & : G \rightarrow \Pow(\bot)\nonumber\\
  & : i_{G} \rightarrow \lbrace c_\bot\ |\ (c_\bot\text{ is contained in }i_{G}) \wedge (c_\bot \in \bot)\rbrace\nonumber
\end{align}
\end{definition}
\vspace{-10pt}
\end{svgraybox}

@@

Thus, the function $cast$ associates a granule $i_G$ in $G$ to a corresponding granule $j_H$ in $H$. Two kinds of granularity-to-granularity mappings can now be discerned: an \emph{upwards mapping} is a mapping from a granularity $G$ to a coarser granularity $H$, whereas a downwards mapping is a mapping from a granularity $K$ to a finner granularity $L$. Orthogonal to this classification, mappings between two granularities may be classified as \emph{regular mappings}, \emph{irregular mappings} or \emph{congruent mappings}~\cite{Lin97},~\cite{DyresonSnodgrass1994}.
\begin{itemize}
\item
\emph{Regular mapping}: A regular mapping is a granularity-to-granularity mapping where the mapping function value is calculated by means of multiplications and/or divisions and (maybe) an anchor adjustment. E.g. the mapping value of the mapping between hours and minutes is calculated using a multiplication by $60$. 
\item
\emph{Irregular mapping}: An irregular mapping is a granularity-to-granularity mapping where the mapping function value can not be calculated by means of multiplications and/or divisions. E.g. the mapping value of the mapping between months and days is dependent on the exact month or day. 
\item
\emph{Congruent mapping}: A congruent mapping is a granularity-to-granularity mapping where the two granularities involved in the mapping have the same granules but a different anchor. E.g. the mapping between the days (Gregorian calendar days) and the academic days is a congruent mapping.
\end{itemize}

@@

\begin{example}
Consider a system that models both Gregorian calendar dates as well as academic calendar dates. In this system, the chronons granularity is a set of milliseconds. Figure \ref{fig:granularity-graph-example} shows the complete granularity graph corresponding to this example. The transitions between the chronons granularity and the seconds granularity and between the seconds granularity and the minutes granularity are examples of regular mappings. They are represented by normal arrows in the visualisation of the graph. The transition between the days granularity and the months granularity is an example of an irregular mapping. In the graph visualisation, irregular mappings are represented by a bold arrow. Finally, the transition between the Gregorian calendar day granularity and the academic day granularity is an example of a congruent mapping. Both concern the same days, but the academic year starts on October 1st, whereas the Gregorian calendar year starts on January 1st. In the graph visualisation, congruent mappings are visualised as straight lines without arrow heads.

@@

\begin{figure}
\centering
\includegraphics[scale=0.5]{graphs/granularityGraph.eps}
\caption{A granularity graph example}
\label{fig:granularity-graph-example}
\end{figure}

@@

\subsection{Temporal Relationships}
%an explanation on allen's temporal relations.
%definition of fuzzy temporal interval
In this section, a brief introduction can be found, concerning \emph{temporal relationships}, sometimes also called `temporal relations'\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}. Temporal relationships can be seen as relationships between temporal elements or instants.

\def\JPicScale{0.5}
\begin{figure}[h]
\centering
\includegraphics[scale=0.5]{graphs/allen.eps}
\caption{Allen relations between two time intervals.}
\label{fig:allen}
\end{figure}

Several operators are defined in order to compare temporal elements or time points. Allen~\cite{Allen83} first described the relationships between time intervals and as a special case, between time points. Figure \ref{fig:allen} shows the temporal relationships Allen discerned. Some proposals can be applied to both crisp and fuzzy temporal intervals~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08},~\cite{garrido2009}.

In the next section, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems.

>>>>

Data models used by information systems (and in specific, temporal database systems) may represent these underlying time domains using \emph{chronons}~\cite{Dyreson1994}, which can informally be discribed as the smallest distinguishable time units that can be used in the system. However, to explain what chronons are, an explanation of some other temporal concepts is necessary.

@@

\begin{definition}\textbf{Duration}\\
A \emph{\textbf{duration}} is an amount of time with known length, but no specific starting or ending instants.
\end{definition}
\vspace{-10pt}
\end{svgraybox}

A time interval as such is bounded by two instants, whereas a duration is not. Also, it should be noted that an instant is in fact a singular case of a time interval.

\begin{svgraybox}
\vspace{-10pt}
\begin{definition}\textbf{Temporal Element}\\
A \emph{\textbf{temporal element}} is a finite union of time intervals.
\end{definition}

@@

@@

Data models for time modelling might now represent an underlying time domain using chronons:

@@

In a system, the granularity with the shortest granules is the \emph{chronons granularity}, which is denoted by `$\bot$'. It is the granularity of which the granules are chronons. 
%It is now possible to define a function $f$, called a \emph{mapping function}, that maps a given granule $i_G$, $i \in \Z$, from a given granularity $G$ to a set of corresponding chronons. The function definition should look like this:

@@

\begin{align}
f & : G \rightarrow \Pow(\bot)\nonumber\\
  &  i_{G} \mapsto \lbrace c_\bot\ |\ (c_\bot\text{ is contained in }i_{G}) \wedge (c_\bot \in \bot)\rbrace\nonumber
\end{align}
\end{definition}
\vspace{-10pt}
\end{svgraybox}

@@

Thus, the function $cast$ associates a granule $i_G$ in $G$ to a corresponding granule $j_H$ in $H$. Two kinds of granularity-to-granularity mappings can now be discerned: an \emph{upwards mapping} is a mapping from a granularity $G$ to a coarser granularity $H$, whereas a downwards mapping is a mapping from a granularity $K$ to a finner granularity $L$. Orthogonal to this classification, mappings between two granularities may be classified as \emph{regular mappings}, \emph{irregular mappings} or \emph{congruent mappings}~\cite{Lin97},~\cite{DyresonSnodgrass1994}.
\begin{itemize}
\item
\emph{Regular mapping}: A regular mapping is a granularity-to-granularity mapping where the mapping function value is calculated by means of multiplications and/or divisions and (maybe) an anchor adjustment. E.g., the mapping value of the mapping between hours and minutes is calculated using a multiplication by $60$. 
\item
\emph{Irregular mapping}: An irregular mapping is a granularity-to-granularity mapping where the mapping function value can not be calculated by means of multiplications and/or divisions. E.g., the mapping value of the mapping between months and days is dependent on the exact month or day. 
\item
\emph{Congruent mapping}: A congruent mapping is a granularity-to-granularity mapping where the two granularities involved in the mapping have the same granules but a different anchor. E.g., the mapping between the days (Gregorian calendar days) and the academic days is a congruent mapping.
\end{itemize}

@@

\begin{example}
\label{examplegrangraph}
Consider a system that models both Gregorian calendar dates as well as academic calendar dates. In this system, the chronons granularity is a set of milliseconds. Figure \ref{fig:granularity-graph-example} shows the complete granularity graph corresponding to this example. The transition between the chronons granularity and the seconds granularity is an example of a regular mapping. Regular mappings are represented by thin arrows in the visualisation of the graph. The transition between the days granularity and the months granularity is an example of an irregular mapping. In the graph visualisation, irregular mappings are represented by a bold arrow. Finally, the transition between the Gregorian calendar day granularity and the academic day granularity is an example of a congruent mapping. Both concern the same days, but the academic year starts on October 1st, whereas the Gregorian calendar year starts on January 1st. In the graph visualisation, congruent mappings are visualised as straight lines without arrow heads.

@@

\begin{figure}
\centering
\includegraphics[scale=0.5]{graphs/granularityGraph.eps}
\caption{The granularity graph corresponding to example \ref{examplegrangraph}}
\label{fig:granularity-graph-example}
\end{figure}

\vspace{-25pt}

@@

\subsection{Temporal Relationships}
%an explanation on allen's temporal relations.
%definition of fuzzy temporal interval
In this section, a brief introduction can be found, concerning \emph{temporal relationships}, sometimes also called `temporal relations'\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}. Temporal relationships can be seen as relationships between temporal elements belonging to the same time domain. These relationships express how the temporal elements are related to one another, with respect to temporal precedence and overlap.

\def\JPicScale{0.5}
\begin{figure}[h]
\centering
\includegraphics[scale=0.5]{graphs/allen.eps}
\caption{Allen relations between two time intervals I and J.}
\label{fig:allen}
\end{figure}

Several (groups of) operators have been proposed in order to compare temporal elements and model the temporal relationships between them. Allen~\cite{Allen83} most notably described such relationships between time intervals and as a special case, between instants. Figure \ref{fig:allen} shows the temporal relationships Allen discerned. Some proposals can be applied to both crisp and other time intervals~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08},~\cite{garrido2009}.

%In the next section, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems.
